---
title: "4. Differences in individual dominance"
author: "Justin Varholick"
date: "2023-06-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)

library(lubridate)
library(igraph)

library(ggplot2)
library(ggsci)
library(cowplot)
library(stringr)
```
##Extract data from BORIS
1 Open project file in BORIS
2 Click the Observations Menu > Export Events > aggregated events
3 Select the observations you wish to export, then all subjects and all behaviors
4 Click YES for group events into one file
5 Save as .csv file
6 edit in excel 
  1. split by "_" delimiter into 6 columns
  2. remove "MVI" from data
  3. Check all dates are actual dates (remove "2020")
  4. check all hours are actual hours (replace pt1, and shift 21 and 12 which extend to next column)
  5. check that all columns have the correct data, delete the coder initials data

##Import data into R
```{r}
adata <- read.csv("/Users/justinvarholick/Documents/GitHub/AcomysDominance_2023/Data/dominanceevents_MASTER_adj.csv", stringsAsFactors=TRUE)

###Fixing observation.id inconsistencies...
adata$Cage <- adata$Observation.id
adata$Subject <- gsub("Mouse ","", as.factor(adata$Subject))
adata$Date <- gsub("-Jan","01", adata$Date)
adata$Modifiers <- gsub("None", "3", adata$Modifiers)
adata$Date <- gsub("3001", "20200130", adata$Date)
adata$Date <- gsub("2901", "20200129", adata$Date)
adata$Date <- gsub("2801", "20200128", adata$Date)
adata$Date <- gsub("2301", "20200123", adata$Date)
adata$Date <- gsub("2201", "20200122", adata$Date)
adata$Date <- gsub("2101", "20200121", adata$Date)
adata$Date <- gsub("730", "20190730", adata$Date)
adata$Date <- gsub("731", "20190731", adata$Date)
adata$Date <- gsub("801", "20190801", adata$Date)
adata$Date <- gsub("806", "20190806", adata$Date)
adata$Date <- gsub("807", "20190807", adata$Date)
adata$Date <- gsub("808", "20190808", adata$Date)
adata$Date <- gsub("815", "20190815", adata$Date)
adata$Date <- gsub("816", "20190816", adata$Date)
adata$Date <- gsub("817", "20190817", adata$Date)
adata$Date <- gsub("2019020190731", "20190731", adata$Date)
adata$Date <- gsub("2019020190807", "20190807", adata$Date)
adata$Date <- gsub("916", "20200916", adata$Date)
adata$Date <- gsub("922", "20200922", adata$Date)
adata$Date <- gsub("923", "20200923", adata$Date)
adata$Date <- gsub("2020020200916", "20200916", adata$Date)
adata$Date <- gsub("2020020200922", "20200922", adata$Date)
adata$Date <- gsub("2020020200923", "20200923", adata$Date)

###Organize data
adata$Hour2 <- paste(adata$Hour, "00", sep=":")
adata$Date <- ymd(adata$Date)
adata$DateHour <- paste(adata$Date, adata$Hour2, sep=" ")
adata$DateHour <- as.POSIXct(adata$DateHour, tz="America/New_York")
adata$DateHour <- ymd_hms(adata$DateHour)

adata$Start..s. <- duration(adata$Start..s., "seconds")
adata$Stop..s. <- duration(adata$Stop..s., "seconds")

adata$Start <- (adata$DateHour + adata$Start..s.)
adata$Stop <- (adata$DateHour + adata$Stop..s.)

ag <- adata %>% 
  within(Opponent.1 <- paste(Cage, Subject, sep="-")) %>% 
  within(value <- paste("1")) %>% 
  select(Opponent.1, value, Date, Behavior, Behavioral.category, Modifiers, Cage, Week)

locations <- ag %>% 
  filter(Behavioral.category=="Location") %>% 
  filter(Behavior!="Shelter") %>% 
  filter(Behavior!="Food Hopper") %>% 
  filter(Behavior!="Cup") %>% 
  unite("Behavior2", Behavior, Modifiers, sep="-", remove=T) %>% 
  rename("Behavior"=Behavior2)

ag <- ag %>% 
  filter(Behavioral.category!="Location") %>% 
  select(Opponent.1, value, Date, Behavior, Behavioral.category, Cage, Week)
ag <- rbind(ag, locations)

###add independent variables
inventory <- read_excel("/Users/justinvarholick/Documents/GitHub/AcomysDominance_2023/Data/AcomysDom_Inventory.xlsx", sheet = "Sheet1")
inventory$Cage <- as.factor(inventory$Cage)
inventory <- inventory %>% 
  select(AnimalID, Trt, Sex, AgeAtWeek1) %>% 
  mutate(Opponent.1=AnimalID)
ag$Cage <- as.factor(ag$Cage)

ag <- left_join(ag, inventory, by="Opponent.1")

ind_ds <- read.csv("/Users/justinvarholick/Documents/GitHub/AcomysDominance_2023/Data/dscorebyanimal_edited.csv")
ind_ds$Cage <- as.factor(ind_ds$Cage)
ind_ds <- rename(ind_ds, "Opponent.1"=AnimalID)

ag <- left_join(ag, ind_ds, by=c("Opponent.1", "Cage"))
ag$value <- as.numeric(ag$value)
```
###Counts per animal, add in zero matrix
```{r}
#data organization
ag2 <- ag %>% 
  group_by(Opponent.1, Cage, Trt, AgeAtWeek1, Behavior) %>% 
  summarise(TotalFrequency=sum(value))

zeromatrix <- ag %>%
  select(Opponent.1, Cage, Trt, AgeAtWeek1, Behavior) %>% 
  distinct() %>% 
  mutate(Chasing=0) %>% 
  mutate(Mounting=0) %>% 
  mutate(Attacking=0) %>% 
  mutate(Stealing=0) %>% 
  mutate(Fleeing=0) %>% 
  mutate(Displace=0) %>%
  mutate(Freeze=0) %>% 
  mutate(Unseen=0) %>% 
  mutate(Inactive=0) %>% 
  mutate(`Top-Left`=0) %>% 
  mutate(`Top-Right`=0) %>% 
  mutate(`Bottom-Left`=0) %>% 
  mutate(`Bottom-Right`=0) %>% 
  gather("Behavior", "n_prime", 6:18) %>% 
  distinct()

ag_zeros <- left_join(zeromatrix, ag2, by=(c("Opponent.1", "Cage", "Trt", "AgeAtWeek1", "Behavior")))
ag_zeros[is.na(ag_zeros)]<-0

ag3 <- ag_zeros %>% 
  filter(Behavior!="Top-Right") %>% 
  filter(Behavior!="Top-Left") %>% 
  filter(Behavior!="Bottom-Right") %>% 
  filter(Behavior!="Bottom-Left") %>% 
  filter(Behavior!="Unseen") %>% 
  filter(Behavior!="Inactive")

ag3$Behavior <- factor(ag3$Behavior, levels=c("Chasing", "Mounting", "Attacking", "Stealing", "Fleeing", "Displace", "Freeze"))
```


##Differences between sexes
###plots
```{r}
ag_sexes <- ag3 %>% 
  filter(Trt =="YFP" | Trt =="YMP")

ggplot(ag_sexes, aes(Behavior, TotalFrequency, color=Trt, fill=Trt))+
  stat_summary(mapping=aes(Behavior, TotalFrequency, fill=Trt), geom="col", fun="mean", color="black", position=position_dodge2(width=0.90))+
  stat_summary(mapping=aes(Behavior, TotalFrequency, group=Trt), geom="errorbar", fun.data="mean_se", color="black", size=0.5, position=position_dodge2(width=0.90))+
  scale_y_continuous(expand=c(0,0))+
  scale_color_npg()+
  scale_fill_npg()+
  ylab("Total freq. per animal") + xlab("")+
  theme_cowplot()+
  theme(legend.position = "right")

```
###stats
```{r}
m.sex.bhv <- lmer(TotalFrequency ~ Behavior + Trt + Behavior*Trt + (1|Cage), data=ag_sexes)
anova(m.sex.bhv)
summary(m.sex.bhv)
```
##Differences between Group sizes for males
###plots
```{r}
ag_grpsize <- ag3 %>% 
  filter(Trt =="YMT" | Trt =="YMP") 

ggplot(ag_grpsize, aes(Behavior, TotalFrequency, color=Trt, fill=Trt))+
  stat_summary(mapping=aes(Behavior, TotalFrequency, fill=Trt), geom="col", fun="mean", color="black", position=position_dodge2(width=0.90))+
  stat_summary(mapping=aes(Behavior, TotalFrequency, group=Trt), geom="errorbar", fun.data="mean_se", color="black", size=0.5, position=position_dodge2(width=0.90))+
  scale_y_continuous(expand=c(0,0))+
  scale_color_manual(values=c("#00A087FF", "#3C5488FF"))+
  scale_fill_manual(values=c("#00A087FF", "#3C5488FF"))+
  ylab("Total freq. per animal") + xlab("")+
  theme_cowplot()+
  theme(legend.position = "right")

```
###stats
```{r}
m.grpsize.bhv <- lmer(TotalFrequency ~ Behavior + Trt + Behavior*Trt + (1|Cage), data=ag_grpsize)
anova(m.grpsize.bhv)
summary(m.grpsize.bhv)
```
##Differences between Ages 
###plots
```{r}
ggplot(ag3, aes(AgeAtWeek1, TotalFrequency))+
  geom_point()+
  geom_smooth(method="lm", se=F)+
  facet_wrap(vars(Behavior))+
  ylab("Total freq. per animal") + xlab("Age in days")+
  theme_cowplot()

ag_age <- ag3 %>% 
  filter(Trt =="YFP" | Trt =="AFP") 

ggplot(ag_age, aes(Behavior, TotalFrequency, color=Trt, fill=Trt))+
  stat_summary(mapping=aes(Behavior, TotalFrequency, fill=Trt), geom="col", fun="mean", color="black", position=position_dodge2(width=0.90))+
  stat_summary(mapping=aes(Behavior, TotalFrequency, group=Trt), geom="errorbar", fun.data="mean_se", color="black", size=0.5, position=position_dodge2(width=0.90))+
  scale_y_continuous(expand=c(0,0))+
  scale_color_manual(values=c("#00A087FF", "#3C5488FF"))+
  scale_fill_manual(values=c("#00A087FF", "#3C5488FF"))+
  ylab("Total freq. per animal") + xlab("")+
  theme_cowplot()+
  theme(legend.position = "right")
```
###stats
```{r}
ag_age_stats <- ag3 %>% 
  group_by(Behavior) %>% 
  summarise(cor=cor.test(TotalFrequency, AgeAtWeek1, method="spearman")$estimate, pval=cor.test(TotalFrequency, AgeAtWeek1, method="spearman")$p.value)

m.age.bhv <- lmer(TotalFrequency ~ Behavior + Trt + Behavior*Trt + (1|Cage), data=ag_age)
anova(m.age.bhv)
summary(m.age.bhv)
```


